---
title: "ADA Homework 2"
author: "Samantha Rabinowitz, sar4357"
date: "3/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Challenge 1

The R code below will load the 'movies.csv' dataset from GitHub and organize the data into a tibble.

```{r}
library(tidyverse)

f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f, col_names = T)
glimpse(d)
```

The following code will filter the dataset to just include movies from 1920 to 1979 and movies that are less than 4 hours long. Columns were also added to make **startYear** a new variable called **decade**.

```{r}
d %>%
  filter(startYear >= "1920" & startYear <="1979" & runtimeMinutes < 240) %>%
  mutate(decade = case_when(startYear >=1920 & startYear<=1929 ~ "20s",
                            startYear >=1930 & startYear<=1939 ~ "30s",
                            startYear >=1940 & startYear<=1949 ~ "40s",
                            startYear >=1950 & startYear<=1959 ~ "50s",
                            startYear >=1960 & startYear<=1969 ~ "60s",
                            startYear >=1970 & startYear<=1979 ~ "70s")) %>%
  glimpse()
```

The code below utilizes *ggplot* to plot histograms of the distribution of **runtimeMinutes** for each decade. 

```{r}
d %>% 
  filter(startYear >= "1920" & startYear <="1979" & runtimeMinutes < 240) %>% 
  mutate(decade = case_when(startYear >=1920 & startYear<=1929 ~ "20s",
                            startYear >=1930 & startYear<=1939 ~ "30s",
                            startYear >=1940 & startYear<=1949 ~ "40s",
                            startYear >=1950 & startYear<=1959 ~ "50s",
                            startYear >=1960 & startYear<=1969 ~ "60s",
                            startYear >=1970 & startYear<=1979 ~ "70s")) %>%
  ggplot(aes(x=runtimeMinutes)) + geom_histogram(bins=60) + facet_wrap(~ decade)
```
