---
title: "ADA Homework 2"
author: "Samantha Rabinowitz, sar4357"
date: "3/3/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Challenge 1

The R code below will load the 'movies.csv' dataset from GitHub and organize the data into a tibble.

```{r, include=FALSE}
library(tidyverse)
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f, col_names = T)
glimpse(d)
```

The following code will filter the dataset to just include movies from 1920 to 1979 and movies that are less than 4 hours long. Columns were also added to make **startYear** a new variable called **decade**.

```{r}
d1 <- d %>%
  filter(startYear >= "1920" & startYear <="1979" & runtimeMinutes < 240) %>%
  mutate(decade = case_when(startYear >=1920 & startYear<=1929 ~ "20s",
                            startYear >=1930 & startYear<=1939 ~ "30s",
                            startYear >=1940 & startYear<=1949 ~ "40s",
                            startYear >=1950 & startYear<=1959 ~ "50s",
                            startYear >=1960 & startYear<=1969 ~ "60s",
                            startYear >=1970 & startYear<=1979 ~ "70s"))
d1 %>% glimpse()
```

The code below utilizes *ggplot* to plot histograms of the distribution of **runtimeMinutes** for each decade. 

```{r}
d1 %>%
  ggplot(aes(x=runtimeMinutes)) + geom_histogram(bins=60) + facet_wrap(~ decade)
```

The R code below will compute the population mean and population standard deviation in **runtimeMinutes** for each decade and store the values in a new dataframe, *results*.

```{r, include=FALSE}
library(radiant)
```

```{r}
results <- d1 %>% group_by(decade) %>% summarize(mean=mean(runtimeMinutes),
                                      pop_sd=sdpop(runtimeMinutes))
results %>% glimpse()
```

The following code will generate a function to calculate the standard error of the mean for each decade as well as a single sample of 100 movies from each decade and calculate the sample mean and standard deviation for each decade. Additionally, the SE around each population mean for each decade will be estimated using the standard deviation and sample size of these samples.

```{r}
std_error <- function(x) {
  sd(x) / sqrt(length(x))
}
d1 %>% group_by(decade) %>% sample_n(100, replace=FALSE) %>% 
  summarize(mean(runtimeMinutes), sd(runtimeMinutes),std_error(runtimeMinutes))
```

The code below will write a function to calculate the standard error of the mean for each deacde using the population standard deviation for purposes of comparison to the values obtained from the sample created above.

```{r}
pop_std_error <- function(x) {
  sdpop(x) / (sqrt(100))
}
d1 %>% group_by(decade) %>%
  summarize(mean = mean(runtimeMinutes),
            sdpop = sdpop(runtimeMinutes),
            pop_se = pop_std_error(runtimeMinutes))
```

The following will generate a sampling distribution of mean **runtimeMinutes** for each decade by (a) drawing 10,000 samples of 100 movies from each decade and, for each sample, (b) calculating the mean **runtimeMinutes** and the standard deviation in **runtimeMinutes**. 

```{r, include=FALSE}
library(mosaic)
```

```{r}
k <- 1000
n <- 100
s <- list(length=k)
  for (i in unique(d1$decade)) {
    d1 %>% filter(i==decade)
    s[[i]] <- (do(k) * sample(d1, size = n, replace = FALSE))
  }
head(s)
```

## Challenge 2

```{r}
ppois(13, lambda = 18)
dpois(0, lambda = 18)
dpois(7, lambda = 18)
1-ppois(20, lambda = 18)
```

The probability that she will hear 13 or fewer calls is 0.1425978. The probability that she will hear no calls is 1.522998e-08. The probability that she will hear exactly 7 calls is 0.00185002. The probability tha she will hear 20 calls or more 0.2692798.

The below code will plot a Poisson mass function with a lambda value of 18 over an x range of 0 to 40.

```{r}
plotDist("pois", lambda = 18, main = "Poisson Distribution\nwith lambda = 18", xlab = "x", xlim = c(0:40), ylab = "Pr(X=x)")
```

The following R code will simulate 520 (10 years of Saturday call omnitoring sessions) results from the above distribution. Following creation of these results, a histogram will be plotted to compare to the shape of the above probability mass function.

```{r}
calls <- rpois(520, lambda = 18)
histogram(calls, xlim=c(0:40))
```

The general shape of the histogram of simulated results above reflects what is seen in the Poisson probability mass function. If the simulation included a greater number of results, this histogram may more closely match what is shown in the probability ass function plot.

## Challenge 3

The R code below will load the 'zombies.csv' dataset from GitHub and organize the data into a tibble.

```{r}
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/zombies.csv"
d <- read_csv(f, col_names = T)
glimpse(d)
```

The code below will calculate the population mean and standard deviation for each quantitative random variable.

```{r}
d %>% summarize(height_mean=mean(height),
                weight_mean=mean(weight),
                age_mean=mean(age),
                n_zombies_mean=mean(zombies_killed),
                ed_mean=mean(years_of_education))
d %>% summarize(height_sd=sdpop(height),
                weight_sd=sdpop(weight),
                age_sd=sdpop(age),
                n_zombies_sd=sdpop(zombies_killed),
                ed_sd=sdpop(years_of_education))
```

The following will utilize *ggplot2* to make scatterplots of height and weight in relation to age. 

```{r}
d %>% ggplot(aes(x=age,y=height)) + geom_point()
d %>% ggplot(aes(x=age,y=weight)) + geom_point()
```

The scatterplot generated for height in relation to age shows a distinct positive relationship. The scatterplot generated for weight in relation to age does not show as strong of a relationship however there is still a positive tendency to the relationship between the two variables.

The following code will generate histograms and Q-Q plots for the numeric variables in the dataset to test for normality of the data.

```{r}
histogram(d$height, main= "Histogram of Height")
qqnorm(d$height, main="QQ Plot of Height")
qqline(d$height)

histogram(d$weight, main= "Histogram of Weight")
qqnorm(d$weight, main="QQ Plot of Weight")
qqline(d$weight)

histogram(d$zombies_killed, main= "Histogram of Zombies Killed")
qqnorm(d$zombies_killed, main="QQ Plot of Zombies Killed")
qqline(d$zombies_killed)

histogram(d$years_of_education, main= "Histogram of Years of Education")
qqnorm(d$years_of_education, main="QQ Plot of Years of Education")
qqline(d$years_of_education)

histogram(d$age, main= "Histogram of Age")
qqnorm(d$age, main="QQ Plot of Age")
qqline(d$age)
```

As demonstrated by the histograms and Q-Q plots above, the height, weight, and age variables seem to be drawn from a normal distribution. The plots generated for the number of zombies killed and years of education indicate that these variables are not drawn from a normal distribution.